How many researchers use 'boilerplate' statistical analysis sections?

### 2021-03-08 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

Many researchers may not have received adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
However, using the wrong statistical methods can cause real harm [@Altman1994] and bad statistical practices are being to used abet weak science [@Stark2018].
Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science [@Allison2016]. Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results [@Ernst2017;@Zhou2018].

@@

As statisticians we have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using 'boilerplate' statistical methods sections.
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

@@

PLOS ONE is a large open access journal that publishes original research across the fundamental sciences. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”. Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results. For example, the EQUATOR network has developed reporting guidelines for study designs commonly used in health research, in part because of the long history of poor statistical application and reporting in health and medical journals [@Altman2016].

All articles published in PLOS ONE are freely accessible via the PLOS Application Programming Interface (API). This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research. Our search strategy comprised of targeted API searches followed by article filtering based on available section headings.

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere within a full-text article. Individual search terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the Material and Methods section versus Supplementary Information. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.

@@

ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting. Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes. As part of study design, researchers are required to provide a summary of statistical methods and planned analyses. Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. All trials registered on AZNCTR are publicly available and can be searched for via an online portal.
[TODO: details of how stat.section text was extracted for analysis]

>>>>

Many researchers lack adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
However, using the wrong statistical methods can cause real harm [@Altman1994] and bad statistical practices are being to used abet weak science [@Stark2018].
Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science [@Allison2016]. Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results [@Ernst2017;@Zhou2018].

@@

Two statisticians on this paper (AB and NW) have heard researchers admit that they sometimes copy-and-paste their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

We used two openly available data sources to find statistical methods sections.

@@

PLOS ONE is a large open access journal that publishes original research across a wide range of scientific fields. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”. Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results. Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].

All articles published in PLOS ONE are freely accessible via the PLOS Application Programming Interface (API). This functionality enabled us to conduct semi-automated searches of full-text articles and analyse data on individual records, including text content and general attributes such as publication date and field(s) of research. To find papers with a statistical methods section we used targeted API searches followed by article filtering based on available section headings.

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in a full-text article. Individual search terms combined the words “data” or “statistical”, with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the Material and Methods section versus Supplementary Information. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.

@@

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting. Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. Details required for registration follow a standardised template (reference or supp file), which covers important details about participant eligibility, intervention(s) being evaluated, study design and outcomes. As part of study design, researchers are required to provide a summary of statistical methods and planned analyses. 
<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->
All trials registered on ANZCTR are publicly available and can be searched for via an online portal.
[TODO: details of how stat.section text was extracted for analysis]
